---
title: "Installing R and R Packages"
author: "Malcolm Barrett"
date: "4/21/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing R

There are several ways to install and use R. Here are two I recommend.

### On the Cloud

1. Go to [rstudio.cloud](https://rstudio.cloud) 
1. Sign up for an account
1. Click on the "Projects" pane
1. Click "Create Project"

This will open up a clean session of Rstudio directly in your browser.

### On your computer

To use R on your computer, you need to download R itself. I also recommend downloading Rstudio, a development environment for R that makes it a lot easier to use.

1. Download and install R from [cloud.r-project.org/](https://cloud.r-project.org/)
1. Download and install Rstudio Desktop from [rstudio.com](https://www.rstudio.com/products/rstudio/download/#download)
1. Open Rstudio (**not R!**)

## Installing R packages

`tidymeta` requires the development version of several packages, including `ggplot2`, to function correctly. You can install the required packages for this vignette with the following code, which you should do even if you have `ggplot2` and friends installed already:


```{r, eval=FALSE}
install.packages(c("devtools", "yaml", "ggrepel"))
library(devtools)
install_github("r-lib/rlang")
install_github("malcolmbarrett/tidymeta")
install_github("malcolmbarrett/mbmisc")
install_github("tidyverse/ggplot2")
```
